---
title: "EPI590R_FINAL"
Name: "Maria Paula Ibarcena Woll"
---

## Packages and Load Dataset File

```{r}
library("gtsummary")
library("here")
library("dplyr")
library("ggplot2")
library("tidyverse")

load((here("Data","cytomegalovirus.rda")),verbose=TRUE)

cytomegalovirus <-cytomegalovirus %>% 
  filter(!is.na(diagnosis.type)) %>%
  mutate(diag_type = case_when(
    diagnosis.type == 1 ~ "Myeloid",
    diagnosis.type == 0 ~ "Lymphoid"))

table(cytomegalovirus$diagnosis.type)

table(cytomegalovirus$diag_type,useNA = "always")


```

## Table 1

```{r}
#| label: tbl-one
#| tbl-cap: "Table One"

## number of observations in the data set
#| eval: true
nrow(cytomegalovirus)

# Create the table 1 and modify it
summary_table <- tbl_summary(
  cytomegalovirus,
  by = diag_type, 
  missing = "no",
  include = c(time.to.transplant, age, sex, race, prior.radiation, prior.transplant, recipient.cmv, donor.cmv, donor.sex),
  label = list(
    time.to.transplant ~ "Time to Transplant",
    prior.radiation ~ "Prior Radiation",
    prior.transplant ~ "Prior Transplant",
    recipient.cmv ~ "CMV Recipient",
    donor.cmv ~ "CMV Donor",
    donor.sex ~ "Donor Sex"
  ),
  missing_text = "Missing"
) %>%
  # Set the title for the entire table
  modify_caption(caption = "**Table 1 - Predictors in DX. Type: Myeloid Vs. Lymphoid**") 

# Print the summary table
summary_table

summary_stat <- inline_text(summary_table, variable = "age", column = "stat_1")
```
In @tbl-one we see the mean and IQR`r summary_stat`


## Logit Regression

```{r}
#| label: tbl-two
#| tbl-cap: "Table two"

# Fit the logistic regression model
logistic_model <- glm(diagnosis.type ~ time.to.transplant + age + sex + race + prior.radiation + prior.transplant + recipient.cmv + donor.cmv + donor.sex,
  data = cytomegalovirus, family = binomial())

# Create a summary table of the logistic model
logistic_model_table <- tbl_regression(
  logistic_model,
  exponentiate = TRUE,  # Exponentiates the coefficients to display as odds ratios
  label = list(
    time.to.transplant ~ "Time to Transplant",
    prior.radiation ~ "Prior Radiation",
    prior.transplant ~ "Prior Transplant",
    recipient.cmv ~ "CMV Recipient",
    donor.cmv ~ "CMV Donor",
    donor.sex ~ "Donor Sex"
  )
)

# View the summary table
logistic_model_table

```
In @tbl-two we can see a binary logistic regression, underscoring the relationship between diagnosis type and time to transplant when controlling for additional covariates including age, sex, race, prior radiation, prior transplant, CMV RECIPIENT, CMV donor and CMV Sex. 

## Boxplot

```{r}
#| label: fig-one
#| fig-cap: "Boxplot of Diagnosis Type and Age"
# Create a box plot 
boxplot<-ggplot(cytomegalovirus, aes(x = diag_type, y = age, fill = diag_type)) +
  geom_boxplot() +
  labs(title = "Box Plot of Age by Diagnosis Type",
       x = "Diagnosis Type",
       y = "Age",
       fill = "Diagnosis Type") +
  theme_minimal() +
  scale_fill_manual(values = c("lightblue", "coral"))
boxplot
ggsave(here("figures","Dx vs Age.png"),
       plot = boxplot)
```
In @fig-one we see a box plot seeing the distribution of age by the Dx type.

## Functions

```{r}
(cytomegalovirus$age)

#calculate mean age in data frame
mean_age <- mean(cytomegalovirus$age, na.rm = TRUE)

# print mean age
print (mean_age)

#calculate sd age in data frame
  sd_age <- sd(cytomegalovirus$age, na.rm = TRUE)
  
# print sd age
print(sd_age)
```
